---
title: "Analyzing Profitability on Short Term Rentals using AirbnB and Zillow Data"
author: "Akshay Kher"
date: "May 2, 2019"
output: html_document
---

# {.tabset .tabset-fade}

## Introduction

#### 1. <span style="color:blue">Question</span>
Help a real-estate agency understand the zip codes that would generate most profit on short term rentals within New York City.

#### 2. <span style="color:blue">Data</span>
The analysis will make use of:

1. **Zillow Data** - Provides the price for two-bedroom properties in NYC
2. **Airbnb Data** - Provides the rent charged for properties in NYC

#### 3. <span style="color:blue">Initial Assumptions</span>
1. The investor will pay for the property in cash (i.e. no mortgage/interest rate will need to be accounted for).
2. The time value of money discount rate is 0% (i.e. $1 today is worth the same 100 years from now).
3. All properties and all square feet within each locale can be assumed to be homogeneous (i.e. a 1000 square foot property in a locale such as Bronx or Manhattan generates twice the revenue and costs twice as much as any other 500 square foot property within that same locale.)


## Setup

### Libraries Required
```{r warning=FALSE, message=FALSE}
require(Rmisc)
require(data.table)
require(DT)
require(tidyverse)
require(stringr)
require(car)
require(kableExtra)
require(ggplot2)
require(plotly)
require(R.utils)
```


### Setup Variables
```{r warnings=FALSE, message=FALSE}
# name of  zillow data file
name_zillow_file <- 'Zip_Zhvi_2bedroom'

# name of airbnb data file
name_airbnb_file <- 'listings' 
```

### Reading data 

```{r warnings=FALSE, message=FALSE}
house_price <- 
  fread(paste0(name_zillow_file, '.csv')) # reading zillow data

house_rent <- 
  fread(paste0(name_airbnb_file, '.csv.gz')) # reading airbnb data
```

## Global Variables

**All code related changes should me made in this section only.**

***

Global variables for house buying price data - Zillow
```{r}
# first date variable in house price data
datetime_start_filter <- "1996-04" 

# last date variable in house price data
datetime_end_filter <- "2017-06"

# city filter
city_filter <- 'New York' 

# first variable in a sequence of factor variables
factor_start_filter <- 'City'

# last variable in a sequence of factor variables
factor_end_filter <- 'CountyName'
```

***

Global variables for house rent price data - Airbnb
```{r}
# subset of variables to be analyzed in house rent data
revenue_vars_filter <- c('zipcode','id', 'last_scraped', 'neighbourhood_group_cleansed',
                         'property_type', 'room_type', 'accommodates', 'bathrooms',
                         'bedrooms', 'beds', 'bed_type', 'square_feet',
                         'security_deposit', 'price', 'cleaning_fee',
                         'guests_included', 'extra_people')

# ASSUMPTION - properties with price greater than $2000 per night (0.01% observations) are removed
price_filter <- 2000

# ASSUMPTION - minimum square feet filter
min_square_feet_filter <- 80 

# ASSUMPTION - maximum square feet filter
max_square_feet_filter <- 10000

# set of numeric columns
numeric_cols_filter <- c('year', 'month', 'day', 'zipcode') 

# set of factor columns
factor_cols_filter <- c('neighbourhood_group_cleansed', 'property_type' ,'room_type') 

# Function to convert price columns to numeric
# Steps:
# 1. store indexes of price columns
# 2. remove the '$' character
# 3. substitute ',' with blank
# 4. convert to numeric
convert_price_to_numeric <- function(data, var)
{
  index <- which(colnames(data) == var) # indexes of price columns
  df1 <- str_sub(data[,index], 2) # remove '$' character
  df2 <- gsub(",","",df1) %>% # substitute ',' with blank
    as.numeric() # convert to numeric
  return(df2)
}

# number of bedrooms filter
filter_bedrooms_filter <- 2 

# ASSUMPTION - years to filter on to calculate the year-on-year price increase
year_filter_subset <- c(2014, 2015, 2016, 2017)

# ASSUMPTION - occupancy rate of house every year
occupancy_rate <- 0.75 

# ASSUMPTION - percentage of stays during which extra people will stay
frequency_extra_people_filter <- 1/5
```

## Preparing Data {.tabset .tabset-fade .tabset-pills}

### Preparing Zillow Data

**Key Data Quality Steps**

1. Date columns have been unpivoted
2. Data types of variables are appropriately converted

***

**Creating a function to prepare the Zillow Data where:**

1. The dataset is tidy:
    * Each variable has its own column
    * Each observation has its own row
    * Each value has its own cell
2. City filter is applied
3. Date columns have been unpivoted
4. RegionName is renamed to zipcode
5. Data types of variables are appropriately converted

```{r}
prepare_zillow_data <- function(house_price,
                                datetime_start=datetime_start_filter,
                                datetime_end=datetime_end_filter,
                                city=city_filter,
                                factor_start=factor_start_filter,
                                factor_end=factor_end_filter)
{
  
  ######### Wrangling Zillow Data ######### 
  # Steps:
  # 1. Select first and last index of date variables
  # 2. Read house price data
  # 3. Filter by city
  # 4. Unpivot date columns
  # 5. Rename RegionName to zipcode
  
    index_datetime_start <- 
    which(colnames(house_price) == datetime_start) # index of first date variable
  
    index_datetime_end <- 
      which(colnames(house_price) == datetime_end) # index of last date variable
    
    house_price_filtered <-
      house_price %>% # read house price data
      filter(City %in% city) %>%  # filter by city
      gather(Datetime, Median_price, index_datetime_start:index_datetime_end) %>%  # unpivot date columns
      separate(Datetime, c("year", "month")) %>%  # separate date into year and month
      rename(zipcode=RegionName) # rename RegionName to zipcode
    
  ######### Changing data types of variables #########
    index_factor_start <- 
      which(colnames(house_price_filtered) == factor_start) # index of first factor variable
    
    index_factor_end <- 
      which(colnames(house_price_filtered) == factor_end) # index of last factor variable
    
    house_price_filtered[,index_factor_start:index_factor_end] <- 
      map(house_price_filtered[,index_factor_start:index_factor_end]
          , as.factor) # converting above set of variables to factor variables
    
    house_price_filtered$year <- 
      as.numeric(house_price_filtered$year) # converting year to numeric
    
    house_price_filtered$month <- 
      as.numeric(house_price_filtered$month) # converting month to numeric
    
  ######### return final data #########
    return(house_price_filtered)
}

```

***

**Using the above function to prepare Zillow Data**
```{r}
house_price_filtered <- prepare_zillow_data(house_price)
```

***

**Zillow Data Dictionary**

```{r zillow, echo = FALSE, message = FALSE, warning = FALSE}
text_tbl <- data.frame (
  Variable = names(house_price_filtered),
  Description = c(
    "Zillow assigned number only, assigned consecutively when the regions are defined",
    "Zip code of where the property is located", 
    "City of where the property is located",
    "State of where the property is located",
    "General name of the surrounding area where the property is located",
    "Political and administrative division of a state, referred to as a particular part of the state",
    "Population of the area; the lower the number the greater the population",
    "Year when data was pulled",
    "Month when data was pulled",
    "Median Price within that area"
  )
)

kable(text_tbl) %>%
  kable_styling(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "30em")
```


***

**Zillow Data**

```{r}
# display first 1000 rows
kable(head(house_price_filtered, 1000)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  scroll_box(width = "100%", height = "500px")
```


### Preparing Airbnb Data


**Key Assumptions**

* Properties with price greater than $2000 per night (0.01% observations) are removed
* Properties with square feet area less than 80 or greater than 10000 are removed
* Occupancy rate of a property each year is 75%
* 20% of the stays will have an extra guest

**Key Data Quality Steps**

1. Removed observations outside the satisfactory square feet range.
2. Filtered dataset for a subset of variables
3. Segregated last_scraped variable into year, month and day
4. Created and applied a function to clean *price* related variables
5. Data types of variables are appropriately converted

**Creating a function to prepare the Airbnb Data where:**

1. The dataset is tidy:
    * Each variable has its own column
    * Each observation has its own row
    * Each value has its own cell
2. Square feet filter is applied
3. Subset of variables filter is applied
4. Data types of variables are appropriately converted

```{r message=FALSE, warning=FALSE}

prepare_airbnb_data <- function(house_rent,
                                revenue_vars=revenue_vars_filter,
                                price_max=price_filter,
                                min_square_feet=min_square_feet_filter,
                                max_square_feet=max_square_feet_filter,
                                numeric_cols=numeric_cols_filter,
                                factor_cols=factor_cols_filter,
                                filter_bedrooms=filter_bedrooms_filter,
                                year_filter=year_filter_subset,
                                occupancy_rate=occupancy_rate_filter,
                                frequency_extra_people=frequency_extra_people_filter)
{
  ######### Wrangling Airbnb Data ######### 
  # Steps:
  # 1. Read house rent data
  # 2. Filter by square feet
  # 3. Filter for subset of variables
  # 4. Separate last_scraped into year, month and day
    house_rent <-
      house_rent %>% # read house rent data
      filter(is.na(square_feet) | square_feet >=min_square_feet, 
             is.na(square_feet) | square_feet <= max_square_feet,  # filter by square feet
             price <= price_max) %>% # filter by price per night
      .[, colnames(.) %in% revenue_vars] %>% # filter by subset of variables
      separate(last_scraped, c('year', 'month', 'day')) # separate last_scraped variable
  
  ######### Changing data types of variables #########
    index_numeric_cols <- 
      colnames(house_rent) %in% numeric_cols # index of numerical columns
    
    index_factor_cols <- 
      colnames(house_rent) %in% factor_cols # index of factor columns
    
    house_rent[,index_numeric_cols] <- 
      map(house_rent[,index_numeric_cols], as.numeric) # convert to numeric
    
    house_rent[,index_factor_cols] <- 
      map(house_rent[,index_factor_cols], as.factor) # convert to factor
    
    # Applying function to convert each price column to numeric
    house_rent$price <- convert_price_to_numeric(house_rent,'price')
    house_rent$security_deposit <- convert_price_to_numeric(house_rent, 'security_deposit')
    house_rent$cleaning_fee <- convert_price_to_numeric(house_rent, 'cleaning_fee')
    house_rent$extra_people <- convert_price_to_numeric(house_rent, 'extra_people')
  
 ######### return final data #########  
  return(house_rent)
}

```

***

**Using the above function to prepare Airbnb Data**
```{r message=FALSE, warning=FALSE}
house_rent <- prepare_airbnb_data(house_rent)
```

***

**Airbnb Data Dictionary**

```{r airbnb, echo = FALSE, message = FALSE, warning = FALSE}
text_tbl <- data.frame (
  Variable = names(house_rent),
  Description = c(
    "Identifier used by AirBnB for the listing",
    "Year when data was pulled", 
    "Month when data was pulled",
    "Day when data was pulled",
    "Name of the area where the property is located",
    "zipcode",
    "Indicates the type of house is this is considered.  Examples are house, apartment, townhomes, etc.",
    "Indicates specific rooms available for rent and/or the entire home",
    "Indicates the amount of people the host is allowing to occupy the property",
    "Indicates the number of bathrooms within the property",
    "Indicates the number of bedrooms within the property",
    "Indicates the number of beds within the property for use",
    "Indicates the type of bed the property has.  ie. futon, real bed, airbed, etc",
    "Square footage of the property or space for rent",
    "Price the host is charging to stay per night",
    "Amount the host would like upfront for any incidentals (should you damage anything during your stay)",
    "Price the host is charging to clean up after your stay",
    "Indicates the number of guests the host is allowing you to bring along",
    "Additional charge per additional guests you bring"
  )
)

kable(text_tbl) %>%
  kable_styling(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "30em")
```

***

**Airbnb Data**

```{r}
# display first 1000 rows
kable(head(house_rent, 1000)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  scroll_box(width = "100%", height = "500px")
```

## Model {.tabset .tabset-fade .tabset-pills}

### Model Building

* **Aim**: Predict the missing square feet of houses

* **Statistial Technique**: Linear Regression

* **Variable Selection Method**: Stepwise AIC

* **Final Variables Selected**:
    * bathrooms
    * price
    * property_type
    * accommodates
    * bedrooms
    
* **Model Adjusted R Square**: 44.53%

***

**Building the model**
```{r results=FALSE}
# Drop rows with specific columns as NAs to avoid error in model building
model_data <- drop_na(house_rent, square_feet, bathrooms, bedrooms) 

# Null Model - Regress square feet on only the intercept
nullmodel=lm(square_feet~1, data=model_data)

# Full Model - Regress square feet on all predictor variables
fullmodel=lm(square_feet~  property_type + room_type + accommodates + 
               bedrooms + bathrooms + beds + price + guests_included + 
               extra_people, data=model_data)

# Final Model built using stepwise variable selection
model_pred_sq_ft <- step(nullmodel, scope=list(lower=nullmodel, upper=fullmodel),
                         direction='both')
```

***

**Pairwise Plots**

```{r}
# square_feet vs bathrooms
p1 <- ggplot(model_data, aes(x=bathrooms, y=square_feet)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("square feet")

# square_feet vs price
p2 <- ggplot(model_data, aes(x=price, y=square_feet)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("square feet")
# square_feet vs accommodates
p3 <- ggplot(model_data, aes(x=accommodates, y=square_feet)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("square feet")

# square_feet vs bedrooms
p4 <- ggplot(model_data, aes(x=bedrooms, y=square_feet)) +
  geom_point(colour = "orange", size = 1.5) +
  geom_smooth(method='lm', color='red') +
  ylab("square feet")

# combining all plots
multiplot(p1, p2, p3, p4, cols=2)
```

***

**Model Summary**
```{r}
summary(model_pred_sq_ft)
```

### Model Diagonostics

1. **Errors are normally distributed with mean=0**

Using a Q-Q Plot, errors are slightly right skewed. As a future step, we will need to diagnose this problem through transforming variables or using a non-linear predictive model.

```{r}
# Constructing a dataframe containing model attributes
model_attributes1 <-
data.frame(index=1:nrow(model_data),
           residuals = model_pred_sq_ft$residuals, 
           fitted_values = model_pred_sq_ft$fitted.values)

# Constructing Q-Q Plot
qqnorm(model_attributes1$residuals)
qqline(model_attributes1$residuals, col='red')
```

***

2. **Uncorrelated Errors **

There seems to be no pattern for the errors over time (index). Thus we can safely assume that the errors are uncorrelated.

```{r}
# Plotting Residuals over Time
model_attributes1 %>%
  ggplot(aes(x=index,y=residuals)) +
  geom_point()
```

***

3. **Constance Variance**

We can see that the residuals have fan-shaped variance across the fitted values. As a future step, we will need to diagnose this problem through transforming variables or using a non-linear predictive model.

```{r}
# Residuals vs Fitted-Value Plot
ggplot(model_attributes1, aes(x=fitted_values,y=residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red") +
  geom_hline(yintercept = 3, color = "blue") +
  geom_hline(yintercept = -3, color = "blue")
```

***

4. **Predictor Variables are independent of each other**

As the **Variation Inflation Factor** < 5 for each predictor variable, we can assume that there is no multi-collinearity.

```{r}
vif(model_pred_sq_ft)
```

***

5. **No influential outliers**

Almost all **standardized errors** are below the absolute value of 4. However, we do have a few standardized errors greater than 4. As a future step, will need to check if it makes sense to remove these observations.

```{r}
# Plotting Studentized/Standardized Errors
rstan <- rstandard(model_pred_sq_ft)  
plot(rstan)
```

## Combining Data

**Creating a function to combine Zillow and Airbnb data: **

1. Both Zillow and Airbnb datasets are filtered to have the same zipcodes
2. **Assumption**: Missing values of bathrooms are filled using the median value
3. **Assumption**: Missing values of square feet are filled using the model built previously
4. **Assumption**: Price per square foot of a property is assumed to be the ratio of median house price and median square feet for a specific neighbourhood
5. **Assumption**: Estimated house Price is assumed to be the product of square feet and price per square foot
6. **Assumption**: Increase in rent and house prices each year are assumed to be equivalent. Further, the year-on-year rent increase post 2017 is calculated as the average house price increase in the last 3 years.
7. **Assumption**: Price per night (2017) of each property is calculated as the sum of:
    * price charged per night
    * price charged for extra people; where we are assuming that extra people will come in 1 out of 5 bookings
8. **Assumption**: Investor is buying the properties in the year 2017.
    
```{r}

combine_data <- function(house_price_data, house_rent_data)
{
  # filter the house rent data using the zipcodes in the house price data
  house_rent_filtered <- 
    filter(house_rent_data, zipcode %in% house_price_data$zipcode
           , bedrooms==filter_bedrooms_filter)
  
  # calculating median bathrooms. 5 NA values for bathrooms will be filled using this
  median_bathrooms <- 
    median(house_rent_filtered$bathrooms, na.rm=TRUE)
  
  # filling missing bathroom values by median value
  house_rent_filtered$bathrooms <- 
    ifelse(is.na(house_rent_filtered$bathrooms), median_bathrooms,
           house_rent_filtered$bathrooms)
  
  # applying model to predict square feet of each property
  house_rent_filtered$predicted_square_ft <- 
    predict(model_pred_sq_ft, newdata = house_rent_filtered)
  
  # filling the missing square feet values by predicted square feet values
  house_rent_filtered$square_feet <- 
    ifelse(is.na(house_rent_filtered$square_feet), 
                  house_rent_filtered$predicted_square_ft, 
                    house_rent_filtered$square)
  
  # remove predicted_square_ft column
  house_rent_filtered$predicted_square_ft <- NULL 
  
  # calculate median house rent by neighbourhood
  median_house_rent_filtered <-
    house_rent_filtered %>% 
    group_by(neighbourhood_group_cleansed) %>% 
    summarize(median_square_feet = median(square_feet, na.rm=TRUE))
  
 ######### combining house rent and house buy data ######### 
 # Steps:
 # 1. add neighbourhood to house price data 
 # 2. calculate median house price by neighbourhood
 # 3. add median house price by neighbourhood to house rent data
 # 4. add median square feet by neighbourhood to house rent data
 # 5. calculate price per square feet
 # 6. calculate buying price of a house as square_feet*price_per_sq_ft
   
  house_rent_and_buy_price <-
    
    # adding neighbourhood to house price data
    inner_join(x = house_price_data,
               y = unique(house_rent_filtered[,c('zipcode', 'neighbourhood_group_cleansed')]),
               by='zipcode') %>% 
    group_by(neighbourhood_group_cleansed, year, month) %>% 
    
    # calculating median house price by neighbourhood
    summarize(median_price = median(Median_price, na.rm=TRUE)) %>% 
    
    # adding median house price by neighbourhood to house rent data
    left_join(x = house_rent_filtered,
              y = .,
              by = c('neighbourhood_group_cleansed','year','month')) %>% 
    
    # adding median square feet by neighbourhood to house rent data
    left_join(median_house_rent_filtered, by='neighbourhood_group_cleansed')
  
  # remove Median_price column
  house_rent_and_buy_price$Median_price <- NULL
  
  # calculating price per square feet
  house_rent_and_buy_price$price_per_sq_ft <- 
    house_rent_and_buy_price$median_price/house_rent_and_buy_price$median_square_feet
  
  # calculating buying price of a house
  house_rent_and_buy_price$house_buying_price <- 
    house_rent_and_buy_price$price_per_sq_ft * house_rent_and_buy_price$square_feet
  
 ######### price per night (2017) for each property ######### 
  
  # calculating percentage increase in house price each year post 2014
  perc_inc_house_price_by_year <- 
    house_price_data %>% 
      group_by(year) %>% 
      summarise(mean_price = mean(Median_price, na.rm = TRUE)) %>% 
      filter(year %in% year_filter_subset) %>% 
      mutate(perc_inc = (mean_price - lag(mean_price))/lag(mean_price))
  
  # averaging year on year house price increase (global)
  approx_perc_inc_rent_price_per_year <<- 
    mean(perc_inc_house_price_by_year$perc_inc, na.rm=TRUE)
  
  # total price per night for each house in the year 2017 
  house_rent_and_buy_price$price_per_night_2017 <- 
    house_rent_and_buy_price$price + frequency_extra_people_filter*house_rent_and_buy_price$extra_people
  
  return(house_rent_and_buy_price)
  }
```
  
***

**Using the above function to prepare Combined Data**

```{r}
house_rent_and_buy_price <- combine_data(house_price_filtered, house_rent)
```

***

**Combined Data Dictionary**

```{r combined, echo = FALSE, message = FALSE, warning = FALSE}
text_tbl <- data.frame (
  Variable = names(house_rent_and_buy_price),
  Description = c(
    "Identifier used by AirBnB for the listing",
    "Year when data was pulled", 
    "Month when data was pulled",
    "Day when data was pulled",
    "Name of the area where the property is located",
    "zipcode",
    "Indicates the type of house is this is considered.  Examples are house, apartment, townhomes, etc.",
    "Indicates specific rooms available for rent and/or the entire home",
    "Indicates the amount of people the host is allowing to occupy the property",
    "Indicates the number of bathrooms within the property",
    "Indicates the number of bedrooms within the property",
    "Indicates the number of beds within the property for use",
    "Indicates the type of bed the property has.  ie. futon, real bed, airbed, etc",
    "Square footage of the property or space for rent",
    "Price the host is charging to stay per night",
    "Amount the host would like upfront for any incidentals (should you damage anything during your stay)",
    "Price the host is charging to clean up after your stay",
    "Indicates the number of guests the host is allowing you to bring along",
    "Additional charge per additional guests you bring",
    "Median house price of the neighbourhood to which the property belongs",
    "Median square foot of the neighbourhood to which the property belongs",
    "Price per square foot",
    "Estimated buying price of the property",
    "Estimated rent charged per night in the year 2017"
  )
)

kable(text_tbl) %>%
  kable_styling(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "30em")
```

***

**Combined Data**
```{r}
# display first 1000 rows
kable(house_rent_and_buy_price) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  scroll_box(width = "100%", height = "500px")
```

## Break-even Time

**The following code calculates the break-even time for each property in years. Break-even time is defined as the number of years it takes to recover the invested money (i.e. money invested to buy the property): **

```{r}

# Steps:
# 1. Run a for loop for each property in the data set
# 2. In each loop:
    # Initalize year = 0. This is the total years needed to breakeven
    # Initalize tot_revenue = 0. This is the cumulative revenue generated through rent
    # rent_price = rent price of the current property for the current year (2017)
    # buy_price = buying price of the current property
  # 3. Run a while loop inside the for loop: Till the buy_price > tot_revenue:
      # increase year by 1
      # increase yearly rent by approx 5.7% (calculated earlier as approx_perc_inc_rent_price_per_year )
      # update tot_revenue: add this yrs revenue to the revenue generated through prev yrs
  # 4. stop when buy_price < tot_revenue:
      # decrease the year by 1 as break-even point has reached before the end of this year
      # calculate revenue needed current year to break-even
      # total time needed to break-even in this year = revenue_req_this_year/yearly rent this year
      # total breakeven time = total time needed this year + previous years
  # 5. Store the breakeven time for the current property in 'breakeven_time' vector

breakeven_time <- c()

for(i in 1:nrow(house_rent_and_buy_price))
{
  year <- 0 # initialize year to 0
  tot_revenue <- 0 # initialize revenue to 0
  rent_price <- 
    house_rent_and_buy_price$price_per_night_2017[i] # rent price of property
  buy_price <- 
    house_rent_and_buy_price$house_buying_price[i] # buy price of property
  
  # while cost of buying the house is greater than the total revenue generated through rent
  while(buy_price>tot_revenue)
  {
    # increase year by 1
    year <- year + 1 
    
    # increase rent price for next year by a fixed percentage
    rent_price <- rent_price * (1+approx_perc_inc_rent_price_per_year) 
    
    # calculating revenue till the current year
    tot_revenue <- tot_revenue + rent_price*365*occupancy_rate
  }
  
  # when the total revenue generated through rent is greater than cost of buying the house 
  
  # decrease the year by 1 as break-even point has reached before the end of this year
  breakeven_year <- year - 1 
  
  # total revenue needed this year to break-even
  revenue_req_this_year <- tot_revenue - buy_price
  
  # total time needed to break-even in this year
  breakeven_year_additional <- revenue_req_this_year/(rent_price*365*occupancy_rate)
  
  # total breakeven time is summation of previous years + time taken in the current year
  breakeven <- breakeven_year + breakeven_year_additional
  
  # storing breakeven time for each property
  breakeven_time[i] <- breakeven
}

# adding breakeven_time variable to the dataset
house_rent_and_buy_price$breakeven_time <- breakeven_time
```

**Data Dictionary**

```{r breakeven, echo = FALSE, message = FALSE, warning = FALSE}
text_tbl <- data.frame (
  Variable = names(house_rent_and_buy_price),
  Description = c(
    "Identifier used by AirBnB for the listing",
    "Year when data was pulled", 
    "Month when data was pulled",
    "Day when data was pulled",
    "Name of the area where the property is located",
    "zipcode",
    "Indicates the type of house is this is considered.  Examples are house, apartment, townhomes, etc.",
    "Indicates specific rooms available for rent and/or the entire home",
    "Indicates the amount of people the host is allowing to occupy the property",
    "Indicates the number of bathrooms within the property",
    "Indicates the number of bedrooms within the property",
    "Indicates the number of beds within the property for use",
    "Indicates the type of bed the property has.  ie. futon, real bed, airbed, etc",
    "Square footage of the property or space for rent",
    "Price the host is charging to stay per night",
    "Amount the host would like upfront for any incidentals (should you damage anything during your stay)",
    "Price the host is charging to clean up after your stay",
    "Indicates the number of guests the host is allowing you to bring along",
    "Additional charge per additional guests you bring",
    "Median house price of the neighbourhood to which the property belongs",
    "Median square foot of the neighbourhood to which the property belongs",
    "Price per square foot",
    "Estimated buying price of the property",
    "Estimated rent charged per night in the year 2017",
    "Time needed (in years) to recover the cost of buying the property"
  )
)

kable(text_tbl) %>%
  kable_styling(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, width = "30em")
```

***

**Final Data**
```{r}
kable(house_rent_and_buy_price) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  scroll_box(width = "100%", height = "500px")
```

## Insights

**Analyzing Cost vs Revenue by Neighbourhood**

* **House Price**
    * The cost of buying a house at *Manhattan > Brooklyn > Queens > Staten Island*
    * Manhattan and Brooklyn have **clear separation** in house prices from other neighbourhoods
    * Queens and Staten Island seem to have **similar house prices**
* **Rent Per Night**
    * The rent per night at *Manhattan > Brooklyn > Queens > Staten Island*
    * Manhattan has **reasonable separation** in rent per night from other neighbourhoods. However, the separation is not as profound as in the case of house prices.
    * Brooklyn, Queens and Staten Island seem to have **similar rent per night**
    
*Conclusion*:\
**A property with high price does not guarantee a high revenue.**\
**Eg**: Brooklyn's property prices are clearly higher than Queens and Staten Island however the rent prices are not proportionally higher. 

Hence we need a more thorough examination of the time it will take to recover the invested money.

```{r warning=FALSE, message=FALSE}
options(scipen = '999')
p5 <- ggplot(house_rent_and_buy_price, 
        aes(x=eval(parse(text='house_buying_price')), 
        fill=eval(parse(text='neighbourhood_group_cleansed')))) +
  geom_histogram(alpha=0.7, position="identity", aes(y = ..density..), color="black") +
  geom_density(alpha=0.7) +
  geom_vline(aes(xintercept=mean(eval(parse(text='house_buying_price'))))
             , color="black", linetype="dashed", size=1) +
  labs(x='House Price', y = "Density") +
  guides(fill=guide_legend(title='Neighbourhood')) + 
  scale_x_continuous(limits = c(0, 6000000))

p6 <- ggplot(house_rent_and_buy_price, 
       aes(x=eval(parse(text='price')), 
           fill=eval(parse(text='neighbourhood_group_cleansed')))) +
  geom_histogram(alpha=0.7, position="identity", aes(y = ..density..), color="black") +
  geom_density(alpha=0.7) +
  geom_vline(aes(xintercept=mean(eval(parse(text='price'))))
             , color="black", linetype="dashed", size=1) +
  labs(x='Price Per Night', y = "Density") +
  scale_x_continuous(limits = c(0, 1000)) +
  guides(fill=guide_legend(title='Neighbourhood'))

multiplot(p5, p6)
```

***

**Analyzing breakeven time and total properties listed by zipcode**

* The breakeven time for *Manhattan > Brooklyn > Queens > Staten Island*
* However, **all zipcodes in Staten Island, except for zipcode *10305*, have 2 or lower properties.** Hence we cannot recommend such zipcodes as they might not represent the true population.
* Zipcodes in Manhattan, Brooklyn and Staten Island have at least 6 properties listed. Hence, we can be more sure of our inference.

```{r warning=FALSE, message=FALSE}
p7 <- ggplot(house_rent_and_buy_price, aes(x = as.factor(zipcode), y = breakeven_time, fill = neighbourhood_group_cleansed)) + 
  geom_boxplot() +
  labs(x = "Zipcode", y = "Breakeven time") + 
  scale_y_continuous(breaks=seq(0, 40, 5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill = guide_legend(title = "Neighbourhood")) +
  labs(fill='Neighbourhood', y='Breakeven Time (years)') 

p8 <-
  ggplot(house_rent_and_buy_price, aes(x=as.factor(zipcode))) +
  geom_histogram(stat = "count", aes(fill = neighbourhood_group_cleansed)) +
  geom_text(stat = "count", aes(label = ..count.., y = ..count.., vjust = -0.2)) +
  labs(x = "Zipcode", y = "Total Listing", fill='Neighbourhood') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, 150, 25))

multiplot(p7, p8)
```

***

**Final Recommendations**

**The final zipcodes recommended are based on how quickly the client can recover the invested money and start earning profit i.e. zipcodes with lower median breakeven time.**

*Note: Zipcodes with 2 or lower properties are not considered*

```{r}
zipcode_recommended <- 
  house_rent_and_buy_price %>% 
  group_by(zipcode, neighbourhood_group_cleansed) %>% 
  summarize(median_breakeven_time = median(breakeven_time),
            count = n()) %>% 
  filter(count > 2) %>% 
  select(-count) %>% 
  arrange(median_breakeven_time) %>% 
  mutate(rank = row_number(), median_breakeven_time = round(median_breakeven_time, 2)) %>% 
  select(rank, zipcode, neighbourhood = neighbourhood_group_cleansed, median_breakeven_time)

zipcode_recommended$rank <- 1:nrow(zipcode_recommended)

datatable(zipcode_recommended)
```

***

**Future Analysis**

* Weigh the revenue generated from each property (after reaching the breakeven point) for providing final recommendations.
* Weigh the review score of each property for providing final recommendations.
* Perform text analytics to understand the sentiment of each review.
* Analyze the impact of amenities on the square feet area.
* Run the same analysis on properties with 1 bhk, 3 bhk, 4 bhk etc. 